Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data

### 2021-07-02 Manuscript/SS_GAM.tex
Revise Method section

\author[a]{Boyi Guo*}
\author[a]{Byron C. Jaeger}
\author[a]{AKM Fazlur Rahman}
\author[a]{D. Leann Long}
\author[a]{Nengjun Yi}

@@

\address[a]{Department of Biostatistics, University of Alabama at
Birmingham, Birmingham, USA}

@@

Semi-parametric regression models are a more appropriate alternative to
the ``black-box'' models, thanks to their balancing between model
flexibility and interpretability. Among those, generalized additive
models (GAMs), proposed in the seminal work of \citet{Hastie1987}, grew
to be one of the most utilized modeling tools. In a GAM, the response
variable, \(Y\), which is assumed to follow some exponential family
distribution \(EF(\mu. \phi)\), can be expressed as the summation of
smooth functions, \(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional
vector of covariates \(\boldsymbol{x}\), written as \[
E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
\] where \(g^{-1}\) is the inverse function of a smooth monotonic link
function. The smooth functions can take many forms and are estimated
using a pseudo-weighted version of the backfitting algorithm
\citep{Breiman1985}. Nevertheless, the classic GAMs cannot fulfill the
increasing analytic demands for high-dimensional data analysis in
biomedical studies.

@@

Following the GLM notation introduced in Section \label{introduction},
we have a generalized additive model \[
E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
\] with the smoothing function \(f_j(x_j)\) of the variable
\(x_j, j = 1, \dots, p.\) The basis function matrix, i.e.~the design
matrix derived from the smoothing function \(f_j(x_j)\), is denoted
\(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
design matrix depends on the choice of the smoothing function. For
example, a cubic spline function with \(k\) knots produces a \(k+2\)
elements column vector. Its corresponding smoothness penalty matrix is
denoted as \(\boldsymbol{S}_j\). In all, the GAMs can be formulated in a
matrix form \[
E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
\] where \(\boldsymbol{\beta}_j\) denotes the basis function
coefficients for the \(j\)th variable, and the matrix transposing
operation is denoted with a superscript \(^t\). To note, the proposed
model can include parametric forms of variables in the model, treating
the parametric function as a special case of the smoothing function,
e.g.~\(f_j(x_j) = x_j\) with the smoothness penalty matrix defined as
\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).

To encourage proper smoothness of the smoothing functions, we adopt the
idea of smoothing penalties from smoothing spline models. The basic idea
is to set up a smoothing penalty, depending on the smoothness penalty
matrix via Equation \ref{eq:smoothpen}, in the posterior density
function, or least squared and likelihood function in the penalized
framework. However, the direct integration of smoothness penalties into
sparsity penalties is not obvious. Given the smoothness penalty matrix
is most often symmetric and positive semi-definite \footnote{ we limit
  the discussion to univariate smoothing function. However, the penalty
  matrix could possibly contain multiple matrices, e.g.~tensor product
  smoother.}, it can be integrated into the design matrix
\(\boldsymbol{X}\) via a re-parameterization. The re-parameterization
follows \citet{Marra2011}, in which eigen-decomposed the penalty matrix
\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\),
where the matrix \(\boldsymbol{D}\) is diagonal with the eigenvalues
arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
\(0\) elements on the diagonal, associated with the null space of the
smooth function. We further write the orthonormal matrix
\(\boldsymbol{U}\), containing the eigenvectors as column, as an (row)
augmentation of two sub-matrices
\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{\text{pen}}\end{bmatrix}\),
where \(\boldsymbol{U}^{\text{pen}}\) contains the eigenvectors (as
columns) for the non-zero eigenvalues, i.e.~the penalized space, and
\(\boldsymbol{U}^0\) contains the eigenvectors for the null space. We
multiply the basis function matrix \(\boldsymbol{X}\) with the
orthonormal matrix \(\boldsymbol{U}\) for the new design matrix
\(\boldsymbol{X}^\ast = \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{X}^0 : \boldsymbol{X}^{\text{pen}} \end{bmatrix}\).
An additional scaling step is imposed on \(\boldsymbol{X}^{\text{pen}}\)
by the non-zero eigenvalues of \(\boldsymbol{D}\) such that the new
basis function matrix \(\boldsymbol{X}^\ast\) is scale-invariant
\footnote{ I know this work for tensor spline (see \citet{Wood2006}),
  but I am not sure if it can be generalized to all smoothing functions}.
For notation convenience, we denote
\(\boldsymbol{X}_j \equiv \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^{\text{pen}} \end{bmatrix}\)
as the after re-parameterized basis function matrix for the \(j\)th
variable hereafter. Similarly, the spline function can be expressed in
the matrix form \[
f_j(x_j) = f_j^0(x_j) + f_j^\text{pen}(x_j) = {\beta_j^0}^T \boldsymbol{X}_j^0 + {\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
\]

@@

\hypertarget{spike-and-slab-spline-prior}{%
\subsection{Spike and Slab Spline
Prior}\label{spike-and-slab-spline-prior}}

For high-dimensional data analysis, a common Bayesian modeling approach
is to impose a particular family of priors, the spike-and-slab (SS)
priors on the coeffcient in a regression modelss. The term
``spike-and-slab'' was first coined in \citet{Mitchell1988} describing
the shape of the prior. It is a mixture distribution, comprised with a
skinny spike density \(f_{spike}(\cdot)\) for weak signals and a flat
slab density \(f_{slab}(\cdot)\) for large signals, mathematically \[
 \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
\]

The most distinct feature of the SS priors family is that it is
conditioned on a latent binary variable \(\gamma_i\) that indicates
whether the covariate \(x_i\) is included in the model. There are
various spike-and-slab priors depending on the choice for the spike
density \(f_{spike}(\cdot)\) and the slab density \(f_{slab}(\cdot)\),
for example \citet{George1993}; \citet{George1997}; \citet{Chipman1996}
for grouped variables; \citet{Brown1998} for multivariate outcomes;
\citet{Ishwaran2005}; \citet{Clyde2004} and reference therein. Here we
introduce the most popular spike-and-slab prior introduced in
\citet{George1993}, as stochastic search variable selection (SSVS).
Given a regression problem \[
Y = \boldsymbol{X}\beta+\epsilon, 
\] we can assume the prior distribution of the regression parameters
\(\beta\) depends on a vector of binary variables
\(\gamma=(\gamma_1, \dots \gamma_p)^\prime\) as follows: given
\(\gamma_i\), \(\beta_i\) are independent distributed as a mixture
priors \[
\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
\] The prior of \(\gamma\) is traditionally assumed to be identically
independent Bernoulli distribution. Even the conjugate normal prior of
\(\boldsymbol{\beta}\) provides analytic simplicity, there is no closed
form solution for the posterior density of \(\boldsymbol{\gamma}\).
\citet{George1993} described a Gibbs sampling algorithm to approximate
the posterior distribution of \(\boldsymbol{\gamma}\) for variable
selection. The same authors later compared SSVS with other Bayesian
variable selection approaches, providing some empirical suggestions on
the advantages and disadvantages of various formulations and estimation
algorithms. \citep{George1997} Among the various problem SSVS addresses,
high-dimensionality is not one of them, due to the intensive computation
needs of MCMC algorithms. \citet{George1997} suggested MCMC algorithms
works well for medium size of predictors (p=25), which is not feasible
for high-dimensional data, where the number of predictors easily exceeds
100.

\hypertarget{spike-and-slab-spline-prior-1}{%
\subsection{Spike and slab spline
prior}\label{spike-and-slab-spline-prior-1}}

A generalized spline model can be formulated as \[
y_i \sim EF(\mu_i, \phi), \quad g(\mu_i) = \sum\limits_{j=1}^p f_j(x_{ij}), \quad i = 1, \dots, n.
\] The link function of the mean can be expressed in a matrix form \[
g(\mu_i) = \sum\limits_{j=1}^p\left[f_j^0(x_{ij}) + f_j^{pen}(x_{ij})\right] = \sum\limits_{j=1}^p\left[{\beta_j^0}^T X_{ij}^0 + {\beta_j^{pen}}^T X_{ij}^{pen}\right]
 = \beta^T X_i,
\] where \[
\begin{aligned}
  X_i = \begin{bmatrix}X^0_{i1}\\X^{pen}_{i1}\\\vdots \\X^0_{ip}\\X^{pen}_{ip}\end{bmatrix}
& &
  \beta = \begin{bmatrix}\beta^0_1\\\beta^{pen}_{1}\\\vdots \\\beta^0_{p}\\\beta^{pen}_{p}\end{bmatrix}
\end{aligned}
\]

\hypertarget{mixture-normal-prior}{%
\subsubsection{Mixture-Normal Prior}\label{mixture-normal-prior}}

Given the full deign matrix
\(X = \begin{pmatrix} X_1 & X_2 & \dots & X_n \end{pmatrix}^T\), the
likelihood function of the generalized model can be written as \[
p(y|X\beta, \phi) = \prod\limits^n_{i=1}p(y_i|\beta^T X_i, \phi).
\]

@@

\hypertarget{mixture-double-exponential-prior}{%
\subsubsection{Mixture Double Exponential
Prior}\label{mixture-double-exponential-prior}}

One of the critics received by the spike-and-slab mixture normal prior
is that the tails of a normal distribution diminishes to zero too fast,
which renders problem when estimating the large effects. In contrast,
distributions with heavier tails are preferred, for example \(t\)
distribution with small degree of freedom and double exponential
distribution. Specifically, double exponential prior is equivalent to
the LASSO penalty in the frequentist models, providing a sparse
solution. Meanwhile, the estimation algorithm of LASSO penalty are
extremely efficient, greatly reducing the computation needs. Recent
developments of spike-and-slab priors include spike-and-slab LASSO
prior, i.e.~spike-and-slab mixture double exponential prior. We extend
the spike-and-slab LASSO prior to high-dimensional GAMs, providing
computationally efficiency and sparse solution to our proposed ssGAM-MN.

The set-up of ssGAM-LASSO model is similar to ssGAM-MN, except we
replace the normal distribution with double exponential distribution.
The prior distributions of the coefficients \(\beta^0_j, \beta^{pen}_j\)
for the \(j\) predictor can be formulated as \[
\begin{aligned}
  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1).
\end{aligned}
\] The hyper-priors for \(\gamma^0_j, \gamma^{pen}_j, \theta_j\) are
kept the same as specified in ssGAM-MN.

\hypertarget{alforithm-for-fitting-ssgams}{%
\subsection{Alforithm for fitting
ssGAMs}\label{alforithm-for-fitting-ssgams}}

@@

@@

>>>>

\author[a]{Boyi Guo*}
\author[b]{Byron C. Jaeger}
\author[a]{AKM Fazlur Rahman}
\author[a]{D. Leann Long}
\author[a]{Nengjun Yi}

@@

\address[a]{Department of Biostatistics, University of Alabama at
Birmingham, Birmingham, USA}
\address[b]{Department of Biostatistics and Data Science, Wake Forest
University, Winston-Salem, USA}

@@

Semi-parametric regression models are a more appropriate alternative to
the ``black-box'' models, thanks to their balancing between model
flexibility and interpretability. Among those, generalized additive
models (GAMs), proposed in the seminal work of \citet{Hastie1987}, grew
to be one of the most utilized modeling tools. In a GAM, the response
variable, \(Y\), which is assumed to follow some exponential family
distribution \(EF(\mu, \phi)\)\footnote{TODO: explain \(\phi\)}, can be
expressed as the summation of smooth functions,
\(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional vector of
covariates \(\boldsymbol{x}\), written as \[
E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
\] where \(g^{-1}\) is the inverse function of a smooth monotonic link
function. The smooth functions can take many forms and are estimated
using a pseudo-weighted version of the backfitting algorithm
\citep{Breiman1985}. Nevertheless, the classic GAMs cannot fulfill the
increasing analytic demands for high-dimensional data analysis in
biomedical studies.

@@

Following the GLM notation introduced in Section \label{introduction},
we have a generalized additive model and its matrix form,
\begin{equation}\label{eq:gam}
E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
\end{equation} with the smoothing function \(f_j(x_j)\) of the variable
\(x_j, j = 1, \dots, p.\) The basis function matrix, i.e.~the design
matrix derived from the smoothing function \(f_j(x_j)\), is denoted
\(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
design matrix depends on the choice of the smoothing function. For
example, a cubic spline function with \(k\) knots produces a \(k+2\)
elements column vector. Its corresponding smoothness penalty matrix is
denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\) denotes the
basis function coefficients for the \(j\)th variable, and the matrix
transposing operation is denoted with a superscript \(^T\). To note, the
proposed model can include parametric forms of variables in the model,
treating the parametric function as a special case of the smoothing
function, e.g.~\(f_j(x_j) = x_j\) with the smoothness penalty matrix
defined as \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).

To encourage proper smoothness of the smoothing functions, we adopt the
idea of smoothing penalties from smoothing spline models. The basic idea
is to set up a smoothing penalty, depending on the smoothness penalty
matrix via Equation \ref{eq:smoothpen}, in the posterior density
function, or least squared and likelihood function in the penalized
framework. However, the direct integration of smoothness penalties into
sparsity penalties is not obvious. Given the smoothness penalty matrix
is most often symmetric and positive semi-definite \footnote{ we limit
  the discussion to univariate smoothing function. However, the penalty
  matrix could possibly contain multiple matrices, e.g.~tensor product
  smoother.}, it can be integrated into the design matrix
\(\boldsymbol{X}\) via a re-parameterization. The re-parameterization
follows \citet{Marra2011}, in which eigen-decomposed the penalty matrix
\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\),
where the matrix \(\boldsymbol{D}\) is diagonal with the eigenvalues
arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
\(0\) elements on the diagonal, associated with the null space of the
smooth function. We further write the orthonormal matrix
\(\boldsymbol{U}\), containing the eigenvectors as column, as an (row)
augmentation of two sub-matrices
\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{\text{pen}}\end{bmatrix}\),
where \(\boldsymbol{U}^{\text{pen}}\) contains the eigenvectors (as
columns) for the non-zero eigenvalues, i.e.~the penalized space, and
\(\boldsymbol{U}^0\) contains the eigenvectors for the null space. We
multiply the basis function matrix \(\boldsymbol{X}\) with the
orthonormal matrix \(\boldsymbol{U}\) for the new design matrix
\(\boldsymbol{X}^\ast = \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{X}^0 : \boldsymbol{X}^{\text{pen}} \end{bmatrix}\).
An additional scaling step is imposed on \(\boldsymbol{X}^{\text{pen}}\)
by the non-zero eigenvalues of \(\boldsymbol{D}\) such that the new
basis function matrix \(\boldsymbol{X}^\ast\) is scale-invariant
\footnote{ I know this work for tensor spline (see \citet{Wood2006}),
  but I am not sure if it can be generalized to all smoothing functions}.
For notation convenience, we denote
\(\boldsymbol{X}_j \equiv \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^{\text{pen}} \end{bmatrix}\)
as the after re-parameterized basis function matrix for the \(j\)th
variable hereafter. Similarly, the spline function can be expressed in
the matrix form \[
f_j(x_j) = f_j^0(x_j) + f_j^\text{pen}(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
\] and the generalized additive model in Equation \ref{eq:gam} be
\begin{equation}\label{eq:gam-repa}
E(Y|\boldsymbol{x}) =  g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} ({\beta_j^0}^T \boldsymbol{X}_j^0 + {\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen})\right].
\end{equation}

@@

\hypertarget{spike-and-slab-smooth-priors}{%
\subsection{Spike-and-Slab Smooth
Priors}\label{spike-and-slab-smooth-priors}}

The family of spike-and-slab (SS) priors regression models dominates
Bayesian high-dimensional analysis for its utility in outcome prediction
and variable selection. We defer to \citet{Bai2021Review} for an
in-depth introduction to spike-and-slab priors. To summary,
spike-and-slab priors are a family of mixture distributions that
comprises a skinny spike density \(f_{spike}(\cdot)\) for weak signals
and a flat slab density \(f_{slab}(\cdot)\) for large signals,
mathematically \[
 \beta|\gamma \sim (1-\gamma)f_{spike}(\beta) + \gamma f_{slab}(\beta).
\] The most distinct feature of the SS priors is that it is conditioned
on a latent binary variable \(\gamma\) that indicates whether the
variable \(x\) is included in the model. There are various
spike-and-slab priors depending on the choice for the spike density
\(f_{spike}(\cdot)\)\footnote{TODO: should I change to notation of \(f\)
  to not confound with the smooth functions?} and the slab density
\(f_{slab}(\cdot)\), for example \citet{George1993}; \citet{George1997};
\citet{Chipman1996} for grouped variables; \citet{Brown1998} for
multivariate outcomes; \citet{Ishwaran2005}; \citet{Clyde2004} and
reference therein.

The major criticism of early spike-and-slab priors receives is being
computationally prohibitive. \citep{Bai2021Review} Since then, many
studies focus on alleviating the computational burden that sampling
algorithms bear, which include EMVS \citep{Rockova2014a}, \footnote{TODO:
  add more citation}. Particularly, the development of Spike-and-slab
Lasso \citep{Rockova2018b, Rockova2018} substantially improves the
scalability of SS models, setting up the theoretical foundation for
generalized models in -omics data analysis \footnote{TODO: add Dr.~Yi's
  papers}. The spike-and-slab lasso prior is composed of two double
exponential distributions with mean 0 and different dispersion
parameters, \(0 < s_0 < s_1\), mathematically,
\begin{equation} \label{eq:ssl}
\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
\end{equation} The SSL also mitigated the problem of EMVS where the weak
signals are not shrink to zero, favored in high-dimensional data
analysis. We noticed that \citet{Bai2021} introduces a spike-and-slab
group prior in the GAM framework, where the densities of the spike and
slab components take the group lasso density \citep{Xu2015}.
\citet{Bai2021} takes an ``all-in-all-out'' strategy for function
selection.

\hypertarget{spike-and-slab-lasso-prior}{%
\subsubsection{Spike-and-Slab Lasso
Prior}\label{spike-and-slab-lasso-prior}}

We introduce a novel prior for GAMs, particularly for high-dimensional
nonlinear effect modeling with bi-level selection. The proposed prior
extends from the spike-and-slab lasso prior described in Equation
\ref{eq:ssl}. Given the reparameterized deign matrix
\(\boldsymbol{X}_j = \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^\text{pen}\end{bmatrix}^T\)
for the \(j\)th variable, we impose a two-part SSL prior to the
coefficients \(\boldsymbol{\beta}_j\). Specifically, we impose
independent group priors on the null space coefficients and on the
penalized space coefficients respectively, \[
\begin{aligned}
  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
  \beta^\text{pen}_{jk} | \gamma^\text{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1), 
\end{aligned}
\] where \(\gamma^0_{j}\in\{0,1\}\) and
\(\gamma^\text{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator
variables, indicating if the model includes the linear effect and the
non-linear effect of the \(j\)th variable respectively. ; \(s_0\) and
\(s_1\) are scale parameters, assuming \(0 < s_0 < s_1\) and given.
These scale parameters \(s_0\) and \(s_1\) can be treated as tuning
parameters and optimized via cross-validation. A discussion of how to
choose the scale parameters comes in Section
\ref{selecting-optimal-scale-values}. To note, this prior differs from
previous group spike-and-slab lasso priors{[}TODO: add Dr.~Yi's
citation{]}, where the \(\boldsymbol{\beta}^0_j\) and
\(\boldsymbol{\beta}^\text{pen}_j\) shares the same indicator variables
\(\gamma_j^0\), \(\gamma_j^\text{pen}\) respectively. It is possible to
relax the assumption such that each coefficient \(\beta\) to have its
own latent indicator \(\gamma\), but at the cost ofcomplicating the
bi-level function selection. Conversely, it is possible to add a more
restrictive assumption on the priors, assuming that one indicator
variable decides the inclusion of both the linear effect and non-linear
effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^{pen}_j\). This
converges\footnote{Is ``becomes'' better?} to the SB-GAM
\citep{Bai2021}.

The reparameterization introduced in Section
\ref{bayesian-hierarchical-additive-models} grants the validity of the
proposed prior. First of all, the smooth function bases are linear
dependent to each other based on the definitions, and ask for extra
attention. The eigeondecomposition remedies the problem. Hence our prior
can be set to be conditionally independent. Secondly, the eigenvalue
scaling provides a panacea to allow unified scale parameters for all
bases of all smoothing functions of variables, where scaling is mandate.

The rest of the hierchical prior follows the traditional spike-and-slab
lasso piror: we set up the hyper-prior of \(\boldsymbol{\gamma}\), to
allow local adaption of the shrinkage, using a bernoulli distribution.
The two indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
\(\gamma^{pen}_j\), shares the same probability parameter, \[
\begin{aligned}
\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
\end{aligned}
\]

This is to leverage the fact that the probability of selecting the bases
of a smooth function should be similar, while allowing different penalty
on the null space and penalty space of the spline function. The hyper
prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
function selection level, while that of \(\gamma_{j}^{pen}\) decides the
smoothness of the spline function at basis function level. Meanwhile, we
specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are independently
distributed for analytic simplicity. We further specify the parameter
\(\theta_j\) follows a beta distribution with given shape parameters
\(a\) and \(b\), \[
\theta_j \sim Beta(a, b).
\] The beta distribution is a conjugate prior for the binomial
distribution and hence provides some computation convenience. For
simplicity, we focus on a special case of beta distribution, uniform
(0,1). When the variable have large effects in any of the bases, the
parameter \(\theta_j\) will be estimated large, which in turn encourages
the model to include the rest of bases, achieving the local adaption
among spline bases. Hereafter, we refer Bayesian hierarchical
generalized additive models with the spike-and-slab lasso prior as the
BHAM-SSL.

\newpage

\hypertarget{other-priors}{%
\subsubsection{Other Priors}\label{other-priors}}

@@

One of the critics received by the spike-and-slab mixture normal prior
is that the tails of a normal distribution diminishes to zero too fast,
which renders problem when estimating the large effects. In contrast,
distributions with heavier tails are preferred, for example \(t\)
distribution with small degree of freedom and double exponential
distribution. Specifically, double exponential prior is equivalent to
the LASSO penalty in the frequentist models, providing a sparse
solution.

\hypertarget{algorithms-for-fitting-bhams}{%
\subsection{Algorithms for Fitting
BHAMs}\label{algorithms-for-fitting-bhams}}

@@

Meanwhile, we specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are
independently distributed for analytic simplicity. However, this set-up
could possibility result in a situation that is not theoretically
possible: the non-linear component is selected, but the linear component
is not. This can be addressed analytically by forcing to include the
linear component when non-linear component is included. Another possible
solution is to impose an dependent structure of \(\gamma_{j}^{pen}\) on
\(\gamma_{j^{0}}\), i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\).

@@

The ``all-in-all-out'' set-up was criticized for its inability to
answering an important research question, if the effects are non-linear
or not. On the other hand, the model can be much more flexible by
allowing each dimension of the penalized space to have an independent
indicator, i.e.~\(\gamma^{pen}_{jk}\) of \(\beta^{pen}_{jk}\) for
\(k = 1\dots, K_j\). This prior set-up generalized to the group
spike-and-slab. The additional flexibility doesn't not necessarily grand
the model neither theoretical justification nor analytic simplicity. The
spline penalty assumes a common smooth shrinkage \(\lambda_j\) applied
to all coefficients of the penalized space, which contradicts to the
model setup. Besides, having multiple indicator variables complicates
the decision of if non-linear effects are necessary.
